Pipeline Performance Figure Generation

### 2022-12-08 TestingPipeline/Scripts/paper_figure_generation.Rmd
Most updated rules.

  if (include_virsorter2) {
    keep_score[input_seqs$max_score>=50] <- keep_score[input_seqs$max_score>=50] + 0.5
    keep_score[input_seqs$max_score>=95] <- keep_score[input_seqs$max_score>=95] + 0.5  
  }

@@

  if (include_deepvirfinder) {
    # remove if DVF calls it not viral
    keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] <- keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] - 1
    # add if DVF calls viral
    keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
    keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
  }

@@

  if (include_tuning_not_viral) {
    # tuning removal 
    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- 0
    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1 
    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- 0
    keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] <- keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] - 0.5 
  }

@@

## differences based on genome size
```{r}
viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)

@@

## visualizing a select set of tools
```{r}
viruses$keep_score_few_tools <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)
viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)
viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = T)
viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)
viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)
```

@@

viruses_long_scores <- viruses %>% 
  select(contains("keep_score_"), size_class, true_virus) %>%
  pivot_longer(cols=contains("keep_score_"), 
               names_to="rule_combination",
               values_to="viral_score") %>% 
  mutate(viral_score=as.factor(round(viral_score))) %>%
  group_by(rule_combination, viral_score, size_class, true_virus) %>%
  summarise(n = n())

@@

```{r}
ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_brewer(palette = "PuOr", ) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~true_virus, scales = "free")
```
### Effect of combining tools
```{r}
viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = T,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_vb_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_vb_dvf_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_vb_dvf_vs2_vs <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = T)

viruses$keep_score_vb_dvf_vs2_vs_tv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = T)

viruses$keep_score_vb_dvf_vs2_vs_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)

@@

viruses_long_scores <- viruses %>% 
  select(contains("keep_score_vb"), size_class, true_virus) %>%
  pivot_longer(cols=contains("keep_score_"), 
               names_to="rule_combination",
               values_to="viral_score") %>% 
  mutate(viral_score=as.factor(round(viral_score))) %>%
  group_by(rule_combination, viral_score, size_class, true_virus) %>%
  summarise(n = n())

@@

```{r}
ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_brewer(palette = "PuOr", ) +
  xlab("") +
  ylab("Number of Sequences") + 
  scale_x_discrete(labels=c("VB", "VB+DVF", "VB+DVF+VS2", "VB+DVF+VS2+VS",
                            "VB+DVF+VS2+VS+addition", "VB+DVF+VS2+VS+addition-removal")) +
  facet_grid(~true_virus, scales = "free")
```

>>>>

  if (include_virsorter2) {
    keep_score[input_seqs$max_score>=0.50] <- keep_score[input_seqs$max_score>=0.50] + 0.5
    keep_score[input_seqs$max_score>=0.95] <- keep_score[input_seqs$max_score>=0.95] + 0.5  
  }

@@

  if (include_deepvirfinder) {
     # add if DVF calls viral
    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
    keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
  }

@@

  if (include_tuning_not_viral) {
    # tuning removal 
    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- 0
    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1 
    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- 0
    keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] <- keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] - 0.5 
    # remove if DVF calls it not viral
    keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] <- keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] - 1

  }

@@

## differences based on genome size
```{r}
viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)

@@

## visualizing a select set of tools
```{r}
viruses$keep_score_high_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)
viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = T)
viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)
viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)
```

@@

viruses_long_scores <- viruses %>% 
  select(contains("keep_score_high"), size_class, true_virus) %>%
  pivot_longer(cols=contains("keep_score_"), 
               names_to="rule_combination",
               values_to="viral_score") %>% 
  mutate(viral_score=as.factor(round(viral_score))) %>%
  group_by(rule_combination, viral_score, size_class, true_virus) %>%
  summarise(n = n())

@@

```{r}
ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_brewer(palette = "PuOr", ) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~true_virus, scales = "free")
```
### Effect of combining tools
```{r}
viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = F,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_dvf_vs <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = F,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = T)

viruses$keep_score_dvf_vs_vb <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = T)

viruses$keep_score_dvf_vs_vb_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = T)

viruses$keep_score_dvf_vs_vb_vs2_tv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = T)

viruses$keep_score_dvf_vs_vb_vs2_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)

@@

viruses_long_scores <- viruses %>% 
  select(contains("keep_score_dvf"), size_class, true_virus) %>%
  pivot_longer(cols=contains("keep_score_"), 
               names_to="rule_combination",
               values_to="viral_score") %>% 
  mutate(viral_score=as.factor(round(viral_score))) %>%
  group_by(rule_combination, viral_score, size_class, true_virus) %>%
  summarise(n = n())

@@

```{r}
ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_brewer(palette = "PuOr", ) +
  xlab("") +
  ylab("Number of Sequences") + 
  scale_x_discrete(labels=c("DVF", "DVF+VS", "DVF+VS+VB", "DVF+VS+VB+VS2",
                            "DVF+VS+VB+VS2+addition", "DVF+VS+VB+VS2+addition-removal")) +
  facet_grid(~true_virus, scales = "free")
```
